Generating Correlated Data for Omics Simulation

### 2024-10-18 manuscript/paper.qmd
Add references and edit manuscript
Use same mouse liver data for comparison to real

We find that CYCLOPS performance depends significantly on the dependence structure of the simulated time series data (@fig-CYCLOPS), indicating that the presence of gene-gene dependence is highly impactful on CYCLOPS' performance.
Overall, CYCLOPS performance is poor in the independent data and improved in both PCA and corpcor data, but surprisingly the Wishart data also shows poor performance.
Since by default CYCLOPS performs dimension reduction to the number of dimensions needed to capture 99% of the variance, we hypothesize that data with dependence allows fewer dimensions to be used which improved the performance.

@@

```{r}
#| label: fig-CYCLOPS
#| fig-cap: "Performance of CYCLOPS on simulated time series data sets based on mouse liver data set (GSE151565, Aaronson et al. 2020). (a) Violin plot comparing absolute circular correlations between true phases and CYCLOPS estimated phases on the simulated data sets under the four dependence methods. (b-e) Examples of CYCLOPS estimated phases. CYCLOPS shows good performance when it separates out points by color (true circadian time)."
cyclops_plot
```

>>>>

We find that CYCLOPS performance depends significantly on the dependence structure of the simulated time series data (@fig-CYCLOPS), indicating that the presence of gene-gene dependence is highly impactful on CYCLOPS' performance.
Overall, CYCLOPS performance is poor in the independent data and improved in both PCA and corpcor data, but surprisingly the Wishart data also shows poor performance.
Since by default CYCLOPS performs dimension reduction to the number of dimensions needed to capture 99% of the variance, this performance difference is likely due to the number of eigengenes needed by CYCLOPS to meet the variance threshold (@fig_CYCLOPS b).

@@

```{r}
#| label: fig-CYCLOPS
#| fig-cap: "Performance of CYCLOPS on simulated time series data sets based on mouse liver data set (GSE151565). (a) Violin plot comparing absolute circular correlations between true phases and CYCLOPS estimated phases on the simulated data sets with the four methods. (b) Violin plot comparing the number of eigengenes used by CYCLOPS between the four methods; the red dotted line indicates the number of eigengenes used by CYCLOPS on the real data (13). (c-f) Examples of CYCLOPS estimated phases on the simulated data sets. CYCLOPS shows good performance when it separates out points by color (true circadian time)."
#| fig-width: 12
#| fig-height: 8
cyclops_plot
```
